---
title: "Applications"
output: html_document
---
## Load Libraries
```{r setup, include=FALSE}
library(rsrHonors)
library(bayesplot)
library(glue)
library(purrr)
library(readr)
library(dplyr)
library(ggplot2)
library(rstanarm)
library(sf)
library(latex2exp)
library(MASS)
library(ggridges)
library(wesanderson)
library(grid)
library(forcats)

```

## Spatial Only

### Simulate Data 
```{r}

# simulate data
set.seed(451)
n_s <- 200


nu <- 2.5
sigma2_sp <- 1
phi_sp <- 0.2
tau2 <- 0.1

x1 <- runif(n_s)
x2 <- runif(n_s)
x3 <- rnorm(n_s)
x4 <- rnorm(n_s)
x5 <- rnorm(n_s)

 
locations <- data.frame(x1 = x1, x2 = x2)


X <- cbind(x3, x4, x5)
beta <- c(1,1,-2)


d <- as.matrix(dist(data.frame(x1 = x1,
                               x2 = x2)))
matern_covariance <- sigma2_sp * (1 + sqrt(5)*d/phi_sp + 5*d^2 / (3*phi_sp^2)) * exp(-1*sqrt(5)*d/phi_sp)


W <- mvrnorm(n = 1, mu = rep(0, n_s), matern_covariance)


y_mean <- 1 + as.numeric(X %*% beta + W)

y_pois <- rpois(n_s, exp(y_mean))
y_norm <- rnorm(n_s, y_mean, sqrt(tau2))
y_binom <- rbinom(n_s, 1, exp(y_mean)/(1 + exp(y_mean)))

sim_data <- data.frame(x1 = x1, x2 = x2, x3 = x3,
                       x4 = x4, x5 = x5,
                       W = rep(W),
                       y_mean = y_mean,
                       y_pois = y_pois,
                       y_norm = y_norm,
                       y_binom = y_binom)

```

### Fit Model
```{r}
linmod <- glm(y_pois ~ x3 + x4 + x5, data = sim_data, family=poisson(link="log")) # find starting values


simulated_sglmm <- poisson_model <- rrp_glm(
        fixed = y_pois ~ x3 + x4 + x5,
        spatial =  ~ x1 + x2,
        data = sim_data,
        family = poisson(),
        covfn = NA,
        iter = 10000,
        chains = 2,
        cores = 1,
        param_start = list("beta" = linmod$coefficients,
                           "s2" = 2,
                           "phi" = 0.5),
        priors = list("beta.normal" = 100, # variance of beta prior
                      "s2.IG" = c(2,2), # inverse gamma params
                      "phi.Unif" = c(0.01, 1.5)), # uniform prior on phi
        tuning = list("beta" = c(sqrt(diag(vcov(linmod)))),
                      "s2" = 0.1,
                      "phi" = 0.01,
                      "w" = 0.1),
        nu = nu,
        rank = 25)
save(simulated_sglmm, file = "first_simulated_sglmm.RData")


mcmc_trace(simulated_sglmm$samples, regex_pars = c("beta", "sigma2"))

mcmc_trace(simulated_sglmm$samples, regex_pars = c("delta_[1-5]\\b"))
```

## Spatio-Temporal

### Simulate Data
Let's start by fitting a simple simulated data set. And this time we'll really tune it. 

```{r}
# simulate data
set.seed(451)
n_s <- 100
n_t <- 50


nu <- 2.5
sigma2_sp <- 1
sigma2_tm <- 1
phi_sp <- 0.2
phi_tm <- 0.2
tau2 <- 0.1

x1 <- rep(runif(n_s), times = n_t)
x2 <- rep(runif(n_s), times = n_t)
x3 <- rnorm(n_s*n_t, sd = 0.05)
x4 <- rnorm(n_s*n_t, sd = 0.05)
x5 <- rnorm(n_s*n_t, sd = 0.05)

time <- 1:n_t / n_t

locations <- data.frame(x1 = x1[1:n_s], x2 = x2[1:n_s])


X <- cbind(x3, x4, x5)
beta <- c(1,1,-1)


d <- as.matrix(dist(data.frame(x1 = x1[1:n_s],
                               x2 = x2[1:n_s])))
matern_covariance <- sigma2_sp * (1 + sqrt(5)*d/phi_sp + 5*d^2 / (3*phi_sp^2)) * exp(-1*sqrt(5)*d/phi_sp)

h <- as.matrix(dist(time))
exp_cov <- sigma2_tm * exp(-(h/phi_tm)^2)

W <- mvrnorm(1, rep(0, n_s), matern_covariance)
V <- mvrnorm(1, rep(0, n_t), exp_cov)


y_mean <- 1 + as.numeric(X %*% beta + rep(W, times = n_t) + rep(V, each = n_s))

y_pois <- rpois(n_s * n_t, exp(y_mean))
y_norm <- rnorm(n_s * n_t, y_mean, sqrt(tau2))
y_binom <- rbinom(n_s * n_t, 1, exp(y_mean)/(1 + exp(y_mean)))

sim_data <- data.frame(x1 = x1, x2 = x2, x3 = x3,
                       x4 = x4, x5 = x5,
                       W = rep(W, times = n_t),
                       V = rep(V, each = n_s),
                       y_mean = y_mean,
                       y_pois = y_pois,
                       y_norm = y_norm,
                       y_binom = y_binom)

```

### Fit GLM

```{r}
set.seed(451)
simulated_glm_pois <- stan_glm(y_pois ~ x3 + x4 + x5, 
         data = sim_data, 
         family = poisson(), iter = 5000, chains = 2)
save(simulated_glm_pois, file = "simulated_glm_pois-3_24_12_54.RData")

simulated_glm_negbin <- stan_glm(y_pois ~ x3 + x4 + x5, 
         data = sim_data, 
         family = neg_binomial_2(), iter = 5000, chains = 2)
save(simulated_glm_negbin, file = "simulated_glm-negbin-3_24_12_54.RData")


ggsave("figures/markov-chains/sim_glm.png", mcmc_trace(simulated_glm_negbin), width = 8, height = 6)
```

```{r}
simulated_glm_pois %>% 
  mcmc_trace()

simulated_glm_pois_predictions <- posterior_predict(simulated_glm_pois)
(sim_glm_pois_ppc <- ppc_dens_overlay(sim_data$y_pois, simulated_glm_pois_predictions) +
  lims(x = c(0, 50)))

ggsave("figures/3-24_sim-glm-pois-ppc.png", sim_glm_pois_ppc, width = 8, height = 6)
simulated_glm_negbin %>% 
  mcmc_trace()

simulated_glm_negbin_predictions <- posterior_predict(simulated_glm_negbin)
mcmc_trace(simulated_glm_negbin)
(sim_glm_negbin_ppc <- ppc_dens_overlay(sim_data$y_pois, simulated_glm_negbin_predictions) +
  lims(x = c(0, 50)))

ggsave("figures/3-24_sim-glm-negbin-ppc.png", sim_glm_negbin_ppc, width = 8, height = 6)
```



### Fit STGLMM
```{r}
set.seed(451)

linmod_nb <- glm.nb(y_pois ~ x3 + x4 + x5, data = sim_data)

linmod <- glm(y_pois ~ x3 + x4 + x5, data = sim_data, family=poisson(link="log")) # find starting values


my_time_start <- Sys.time()
  simulated_stglmm <- stglmm(
    fixed = y_pois ~ x3 + x4 + x5,
    data = sim_data,
    locations = locations,
    times = time,
    family = poisson(),
    covfn = 2.5,
    iter = 30000,
    chains = 2,
    cores = 1,
    param_start = list("beta" = linmod$coefficients,
                       "s2_sp" = 3,
                       "phi_sp" = 0.5,
                       "s2_tm" = 1,
                       "phi_tm" = 0.5),
    priors = list("beta.normal" = 100, # variance of beta prior
                  "s2_sp_IG" = c(2,2), # inverse gamma params
                  "s2_tm_IG" = c(2,2),
                  "phi_sp_unif" = c(0.01, 1.5),
                  "phi_tm_unif" = c(0.01, 1.5)), # uniform prior on phi
    tuning = list("beta" = c(sqrt(diag(vcov(linmod)))[2]/4, sqrt(diag(vcov(linmod)))[2:4]),
                  "s2_sp" = 1.5,
                  "phi_sp" = 0.02,
                  "s2_tm" = 1.5,
                  "phi_tm" = 0.01,
                  "delta" = 0.03,
                  "alpha" = 0.02),
    nu = nu,
    rank_sp = 50,
    rank_tm = 10)
  my_time_end <- Sys.time()


save(simulated_stglmm, file = "simulated_stglmm-4_09.RData")
```

```{r}
load("simulated_stglmm-3_24_10_54.RData")
```


#### Display Diagnostics

```{r}
simulated_stglmm$accepts
(sim_stglmm_beta_trace <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "beta.*" ))

ggsave("figures/markov-chains/sim_stglmm_beta.png", sim_stglmm_beta_trace, width = 8, height = 6)


mcmc_trace(simulated_stglmm$samples,
           regex_pars = "sigma.*" )


(sim_stglmm_hyper_trace <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "(sigma.*)|(phi.*)"))

ggsave("figures/markov-chains/sim_stglmm_hyper.png", sim_stglmm_hyper_trace, width = 8, height = 6)




(sim_stglmm_hyper_trace <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "(sigma.*)|(phi.*)" ))

ggsave("figures/markov-chains/sim_stglmm_hyper.png", sim_stglmm_hyper_trace, width = 8, height = 6)

(sim_stglmm_delta_trace_1 <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "delta_(([1]{0,1}[0-9]\\b)|(2[0-5]\\b))" ))

ggsave("figures/markov-chains/sim_stglmm_delta_1.png", sim_stglmm_delta_trace_1, width = 8, height = 6)

(sim_stglmm_delta_trace_2 <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "delta_(([345][0-9]\\b)|(2[6-9]\\b))" ))

ggsave("figures/markov-chains/sim_stglmm_delta_2.png", sim_stglmm_delta_trace_2, width = 8, height = 6)

(sim_stglmm_alpha_trace <- mcmc_trace(simulated_stglmm$samples,
           regex_pars = "alpha_.*" ))

ggsave("figures/markov-chains/sim_stglmm_alpha.png", sim_stglmm_alpha_trace, width = 8, height = 6)


```

Simulation study has a different meaning, do it for 1000 times or so and check for each of those 1 thousand. 

#### Check Predictions

```{r}

predict_stglmm_sim <- function(location_index, time_index, x3, x4, x5) {
  mean_vals <- as.numeric(simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,"beta_0"] +
              simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,"beta_1"] * x3 +
              simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,"beta_2"] * x4 +
              simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,"beta_3"] * x5 +

              simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,glue("w_{location_index}")] +
              simulated_stglmm$samples[(nrow(simulated_stglmm$samples) - 5000):nrow(simulated_stglmm$samples),,glue("v_{time_index}")])
  rpois(length(mean_vals), exp(mean_vals))
}



posterior_predictions_sim <- as.matrix(pmap_dfc(list("location_index" = rep(1:n_s, times = n_t),
             "time_index" = rep(1:n_t, each = n_s),
             "x3" = sim_data$x3,
             "x4" = sim_data$x4,
             "x5" = sim_data$x5),
        predict_stglmm_sim))

sim_stglmm_ppc <- ppc_dens_overlay(sim_data$y_pois, posterior_predictions_sim) +
  xlim(0, 50)

ggsave("figures/3-24_sim-stglmm-ppc.png", sim_stglmm_ppc, width = 8, height = 6)

```


## Simulation Coverage

Load the data

```{r}
load("simulations/simulations_phi_0.2.RData")
sim_phi_0_2 <- sim_runs
load("simulations/simulations_phi_0.5.RData")
sim_phi_0_5 <- sim_runs
load("simulations/simulations_phi_1.RData")
sim_phi_1_0 <- sim_runs

sim_list <- list(phi_0.2 = sim_phi_0_2,
     phi_0.5 = sim_phi_0_5,
     phi_1.0 = sim_phi_1_0)
```

### Phi = 0.2

```{r}

prepare_sim_table <- function(sim_result) {
  sim_result %>% 
    map_depth(1, bind_rows, .id = "model") %>% 
    bind_rows(.id = "run") %>% 
    rename(quant_05 = "5%",
           quant_25 = "25%",
           quant_50 = "50%",
           quant_75 = "75%",
           quant_95 = "95%") %>% 
    mutate(variable = variable) %>% 
    mutate(variable_tex = factor(variable, levels = c("beta_0",  
                                                      "beta_1",
                                                      "beta_2", 
                                                      "beta_3",
                                                      "phi_sp",
                                                      "phi_tm",
                                                      "sigma2_sp",
                                                      "sigma2_tm"),
                                 labels = c(TeX("$\\beta_0$"),
                                            TeX("$\\beta_1$"),
                                            TeX("$\\beta_2$"),
                                            TeX("$\\beta_3$"),
                                            TeX("$\\phi$"),
                                            TeX("$\\psi$"),
                                            TeX("$\\sigma^2$"),
                                            TeX("$\\theta^2$"))))
}

sim_table_list <- map(sim_list, prepare_sim_table)
```


```{r}
darjeeling <- wes_palette("Darjeeling2", type = "discrete", n = 4)
darjeeling <- c(darjeeling[1], darjeeling[4]) 


plot_point_est <- function(sim_table, phi) {
  vline_data = tibble(xintercept = c(1,1,1,-1, 1, 1,phi, phi),
                    variable_tex = unique(sim_table$variable_tex))

  beta_est <- sim_table %>% 
    filter(as.integer(variable_tex) <= 4) %>% 
    ggplot(aes(y = model)) +
    geom_density_ridges(aes(x = quant_50, fill = model), 
                        color = "grey22") +
    theme_minimal() +
    facet_wrap(~variable_tex, ncol = 2, scales = "free_x",
               labeller = label_parsed) +
    scale_y_discrete(labels = c("GLM", "STGLMM")) +
    theme(legend.position = "none") +
    scale_fill_manual(values = darjeeling) +
    geom_vline(data = filter(vline_data, as.integer(variable_tex) <= 4),
               aes(xintercept = xintercept)) +
    labs(y = NULL,
         x = "Point Estimate (Median)")
  
  ggsave(glue("figures/coverage/beta_point_est_{phi}.png"), beta_est,
         width = 8, height = 6)
  
  

  not_beta_est <- sim_table %>% 
    filter(as.integer(variable_tex) > 4) %>% 
    ggplot() +
    geom_density(aes(x = quant_50), 
                 color = "grey22",
                 fill = darjeeling[2]) +
    theme_minimal() +
    facet_wrap(~variable_tex, ncol = 2, scales = "free",
               labeller = label_parsed) +
    theme(legend.position = "none") +
    geom_vline(data = filter(vline_data, as.integer(variable_tex) > 4),
               aes(xintercept = xintercept)) +
    labs(y = NULL,
         x = "Point Estimate (Median)")

  ggsave(glue("figures/coverage/other_point_est_{phi}.png"), not_beta_est,
         width = 8, height = 6)  
}

map2(sim_table_list, c(0.2,0.5,1), plot_point_est)

```

```{r}
plot_quantiles <- function(n, truth, sim_table) {
  sim_table %>% 
    group_by(model, variable_tex) %>% 
    filter(as.integer(variable_tex) == n) %>% 
    arrange(desc(quant_50)) %>% 
    mutate(index = 1:n()) %>% 
    ungroup() %>% 
    mutate(model = factor(model, levels = c("stglmm_params_quantiles", "glm_params_quantiles"))) %>% 
    ggplot() +
    geom_boxplot(aes(x = index,
                     ymin = quant_05,
                     lower = quant_25,
                     middle = quant_50,
                     upper = quant_75,
                     ymax = quant_95,
                     group = index,
                     fill = model),
                 color = "grey30",
                 stat = "identity") +
    facet_wrap(~model, labeller = labeller(model = c(stglmm_params_quantiles = "STGLMM",
                                                     glm_params_quantiles = "GLM")),
               ncol = 1) +
    geom_hline(yintercept = truth) +
    theme_minimal() +
    scale_fill_manual(values = rev(darjeeling)) +
    annotate("text", label = TeX(glue("True $\\beta_{n-1}$")), x = 110, y = truth) +
    coord_cartesian(xlim = c(0, 100), clip = "off") +
    theme(legend.position = "none",
          plot.margin = unit(c(1,3,1,1), "lines")) +
    labs(x = "Simulation",
         y = "Parameter Estimate")
}

plot_quantiles(1, 1, sim_table_list[[1]])

save_coverage_plots <- function(sim_table, phi) {
  beta_0_plot <- plot_quantiles(1, 1, sim_table)
  ggsave(glue("figures/coverage/beta_0_plot_phi_{phi}.png"), beta_0_plot,
         width = 8, height = 6)
  beta_1_plot <- plot_quantiles(2, 1, sim_table)
  ggsave(glue("figures/coverage/beta_1_plot_phi_{phi}.png"), beta_1_plot,
         width = 8, height = 6)
  beta_2_plot <- plot_quantiles(3, 1, sim_table)
  ggsave(glue("figures/coverage/beta_2_plot_phi_{phi}.png"), beta_2_plot,
         width = 8, height = 6)
  beta_3_plot <- plot_quantiles(4, -1, sim_table)
  ggsave(glue("figures/coverage/beta_3_plot_phi_{phi}.png"), beta_3_plot,
         width = 8, height = 6)
}
save_coverage_plots(sim_table_list[[1]], 0.2)

map2(sim_table_list, c(0.2, 0.5, 1.0), save_coverage_plots)
```

Make table of coverage

```{r}

make_coverage_table <- function(sim_table, phi) {
  truth_table <- tibble(variable = c("beta_0", "beta_1", "beta_2", "beta_3", "phi_sp", "phi_tm", "sigma2_sp", "sigma2_tm"),
                        truth = c(1,1,1,-1,phi,phi,1,1))
  
  sim_table %>% 
    left_join(truth_table, by = "variable") %>% 
    mutate(in_ci_50 = quant_25 < truth & truth < quant_75,
           in_ci_90 = quant_05 < truth & truth < quant_95) %>% 
    group_by(variable, variable_tex, model) %>% 
    summarize(prop_in_ci_50 = mean(in_ci_50),
              prop_in_ci_90 = mean(in_ci_90))
}

map2(sim_table_list, c(0.2, 0.5, 1.0), 
     ~{make_coverage_table(.x, .y) %>% 
         ungroup() %>%  
         dplyr::select(-variable_tex) %>% 
         pivot_longer(starts_with("prop"), names_to = "ci", values_to = "prop") %>% 
         pivot_wider(names_from = variable, values_from = prop)
       })

temp_cov_table <- make_coverage_table(sim_table_list[[1]], 0.2)


# 
# temp_cov_table %>% 
#   ggplot(aes(x = prop_in_ci_50, y = variable, fill = model)) +
#   geom_col(position = "dodge") +
#   theme_minimal() +
#   scale_fill_manual(values = darjeeling, labels = c("GLM", "STGLMM")) +
#   geom_vline(xintercept = 0.5) +
#   scale_y_discrete(labels = c("beta_0" = parse(text = TeX("$\\beta_0$")),
#                               "beta_1" = parse(text = TeX("$\\beta_1$")),
#                               "beta_2" = parse(text = TeX("$\\beta_2$")),
#                               "beta_3" = parse(text = TeX("$\\beta_3$")),
#                               "phi_sp" = parse(text = TeX("$\\phi$")),
#                               "phi_tm" = parse(text = TeX("$\\psi$")),
#                               "sigma2_sp" = parse(text = TeX("$\\sigma^2$")),
#                               "sigma2_tm" = parse(text = TeX("$\\theta^2$")))) +
#   labs(x = "Proportion the True Parameter Value is in the 50% CI",
#        y = NULL,
#        fill = "Model:") + 
#   theme(legend.position = "bottom")
```


## Bird Data

### Load Data


```{r}
house_finch <- read_csv("../../Data/house_finch.csv")
carolina_wren <- read_csv("../../Data/carolina_wren.csv")
```


```{r}
routes <- read_csv("../../Data/Routes/routes.csv", col_types = "dddcdddddddd")
ne_10m <- st_read("../../Data/ne_50m_admin_1_states_provinces_lakes/ne_50m_admin_1_states_provinces_lakes.shp") %>% 
  filter(gu_a3 == "USA") %>% 
  st_transform(crs = 5070) %>% 
  filter(!(name %in% c("Alaska", "Hawaii")))
```

```{r}
house_finch_sp <- house_finch %>% 
  left_join(routes, by = c("countrynum", "statenum", "Route"))

carolina_wren_sp <- carolina_wren %>% 
  left_join(routes, by = c("countrynum", "statenum", "Route")) %>% 
  mutate(Stratum = as.character(Stratum))
house_finch %>% 
  ggplot(aes(x = Year)) +
  geom_histogram(binwidth = 1, boundary = 0)

start = 1990
end = 2010

carolina_wren %>% 
  filter(Year >= start, Year <= end) %>% 
  group_by(statenum, Route) %>% 
  filter(n() == end - start + 1)

house_finch_sp %>% 
  filter(Year >= start, Year <= end) %>% 
  group_by(statenum, Route) %>% 
  filter(n() == end - start + 1)

model_data <- carolina_wren_sp %>% 
  filter(Year >= start, Year <= end) %>% 
  group_by(statenum, Route) %>% 
  filter(n() == end - start + 1) %>% 
  ungroup() %>% 
  mutate(times_norm = (Year - min(Year))/20)
```

### Map Data

```{r}

(map_1990 <- model_data %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  filter(Year == 1990) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(crs = 5070) %>%
  ggplot() +
  geom_sf(data = ne_10m, fill = NA, color = "gray40") +
  geom_sf(aes(size = SpeciesTotal,
              color = SpeciesTotal), 
          alpha = 0.5) +
  scale_color_viridis_c(breaks = c(25, 75, 125), 
                        limits = c(0, 150), alpha = 0.8) +
  scale_size(range = c(0.5, 10), 
             breaks = c(25, 75, 125), 
             limits = c(0, 150),
             guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(color = "Observed Birds",
       size = "Observed Birds"))

(map_2010 <- model_data %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(crs = 5070) %>% 
  filter(Year == 2010) %>% 
  ggplot() +
  geom_sf(data = ne_10m, fill = NA, color = "gray40") +
  geom_sf(aes(size = SpeciesTotal,
              color = SpeciesTotal), 
          alpha = 0.5) +
  scale_color_viridis_c(breaks = c(25, 75, 125), 
                        limits = c(0, 150), alpha = 0.8) +
  scale_size(range = c(0.5, 10), 
             breaks = c(25, 75, 125), 
             limits = c(0, 150),
             guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(color = "Observed Birds",
       size = "Observed Birds"))

ggsave("figures/03-19_map-1990.png", map_1990, width = 8, height = 6)
ggsave("figures/03-19_map-2010.png", map_2010, width = 8, height = 6)


(map_1991 <- model_data %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(crs = 5070) %>% 
  filter(Year == 1991) %>% 
  ggplot() +
  geom_sf(data = ne_10m, fill = NA, color = "gray40") +
  geom_sf(aes(size = SpeciesTotal,
              color = SpeciesTotal), 
          alpha = 0.5) +
  scale_color_viridis_c(breaks = c(25, 75, 125), 
                        limits = c(0, 150), alpha = 0.8) +
  scale_size(range = c(0.5, 10), 
             breaks = c(25, 75, 125), 
             limits = c(0, 150),
             guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(color = "Observed Birds",
       size = "Observed Birds"))
ggsave("figures/03-19_map-1991.png", map_2010, width = 8, height = 6)

```

### Fit GLM

```{r}
set.seed(451)
bird_glm <- stan_glm.nb(SpeciesTotal ~ Stratum, 
         data = model_data, iter = 10000, chains = 2)
getwd()
save(bird_glm, file = "bird_glm_nb.RData")
```



```{r}
load("bird_glm.RData")
```


#### Evaluate Chains

```{r}

mcmc_trace(bird_glm)

ggsave("figures/markov-chains/bird_glm.png", mcmc_trace(bird_glm), width = 8, height = 6)
```

```{r}
bird_glm_predictions <- posterior_predict(bird_glm)

bird_glm_ppc <- ppc_dens_overlay(model_data$SpeciesTotal, bird_glm_predictions) +
  lims(x = c(0, 150)) +
  theme_minimal()

bird_glm_ppc


ggsave("figures/4-7_bird-glm-nb-ppc.png", bird_glm_ppc, width = 8, height = 6)
```

Look at variable distributions of Stratum3: 
Coastal Flatwoods (15 locations, 21 years = 315 observations)
```{r}
bird_glm

bird_glm_stratum3 <- bird_glm %>% 
  mcmc_areas(pars = "Stratum3", 
             prob = 0.95, 
             point_est = "median") +
  theme_minimal() +
  theme(axis.text.y = element_blank())

ggsave("figures/3-20_bird-glm-stratum3.png", bird_glm_stratum3, width = 8, height = 6)

```


### Fit STGLMM

```{r}
set.seed(451)
linmod <- glm(SpeciesTotal ~ Stratum, data = model_data, family=poisson(link="log")) # find starting values

locations <- data.frame(lat = model_data$Latitude, 
                        long = model_data$Longitude) %>% 
  distinct()

bird_stglmm <- stglmm(
  fixed = SpeciesTotal ~ Stratum,
  data = model_data,
  locations = locations,
  times = unique(model_data$Year), #normalize to 0 and 1 so psi doesnt have to be crazy
  family = poisson(),
  covfn = 2.5,
  iter = 30000,
  chains = 2,
  cores = 1,
  param_start = list("beta" = linmod$coefficients,
                     "s2_sp" = 1,
                     "phi_sp" = 0.5,
                     "s2_tm" = 1,
                     "phi_tm" = 0.5),
  priors = list("beta.normal" = 100, # variance of beta prior
                "s2_sp_IG" = c(2,2), # inverse gamma params
                "s2_tm_IG" = c(2,2),
                "phi_sp_unif" = c(0.01, 1.5),
                "phi_tm_unif" = c(0.01, 1.5)), # uniform prior on phi
  tuning = list("beta" = c(sqrt(diag(vcov(linmod)))/2),
                "s2_sp" = 0.1,
                "phi_sp" = 0.05,
                "s2_tm" = 0.4,
                "phi_tm" = 0.05,
                "delta" = 0.05,
                "alpha" = 0.05),
  nu = 2.5,
  rank_sp = 50,
  rank_tm = 10)

save(bird_stglmm, file = "bird_stglmm.RData")
```

#### Evaluate Chains


```{r}
load("bird_stglmm.RData")

bird_stglmm$acceptw

bird_stglmm$accepts
(bird_stglmm_beta_trace <- mcmc_trace(bird_stglmm$samples,
           regex_pars = "beta_[0-9]+\\b" ))

ggsave("figures/markov-chains/bird_stglmm_beta.png", bird_stglmm_beta_trace, width = 8, height = 6)

bird_stglmm_hyper_trace <- mcmc_trace(bird_stglmm$samples,
           regex_pars = "(sigma.*)|(phi.*)" )

ggsave("figures/markov-chains/bird_stglmm_hyper.png", bird_stglmm_hyper_trace, width = 8, height = 6)


(bird_stglmm_delta_trace_2 <- mcmc_trace(bird_stglmm$samples,
           regex_pars = "delta_(([345][0-9]\\b)|(2[6-9]\\b))" ))

ggsave("figures/markov-chains/bird_stglmm_delta_2.png", bird_stglmm_delta_trace_2, width = 8, height = 6)

(bird_stglmm_alpha_trace <- mcmc_trace(bird_stglmm$samples,
           regex_pars = "alpha_.*" ))

ggsave("figures/markov-chains/bird_stglmm_alpha.png", bird_stglmm_alpha_trace, width = 8, height = 6)
```

```{r}
model_data_index <- model_data %>% 
  group_by(statenum, Route) %>% 
  mutate(location_index = group_indices(),
         time_index = 1:n())

model_data_minimal <- model_data_index %>% 
  ungroup() %>% 
  dplyr::select(SpeciesTotal, location_index, Stratum, time_index)

# beta for each observation
```

#### Evaluate Results


```{r}
posterior_predictions <- as.vector(bird_stglmm$samples[,,"beta_0"] + bird_stglmm$samples[,,"beta_3"] + bird_stglmm$samples[,,glue("w_{1}")] + bird_stglmm$samples[,,glue("v_{1}")])


predict_stglmm <- function(location_index, Stratum_index, time_index) {
  mean_pred <- as.vector(bird_stglmm$samples[28000:30000,,"beta_0"] +
              bird_stglmm$samples[28000:30000,,glue("beta_{Stratum_index}")] +
              bird_stglmm$samples[28000:30000,,glue("w_{location_index}")] +
              bird_stglmm$samples[28000:30000,,glue("v_{time_index}")])
  
  rpois(length(mean_pred), exp(mean_pred))
}

posterior_predictions <- as.matrix(pmap_dfc(list("location_index" = model_data_minimal$location_index,
             "Stratum_index" = as.numeric(as.factor(paste0("Stratum", model_data_minimal$Stratum)))-1,
             "time_index" = model_data_minimal$time_index),
        predict_stglmm))



bird_stglmm_ppc <- ppc_dens_overlay(model_data$SpeciesTotal, posterior_predictions) +
  lims(x = c(0, 150)) +
  theme_minimal()

bird_stglmm_ppc

ggsave("figures/4-09_bird-stglmm-ppc.png", bird_stglmm_ppc, width = 8, height = 6)
```

Lets look at Stratum3 for this version too
```{r}
stratum3_estimates <- as.vector(bird_stglmm$samples[5001:10000,,"beta_11"]) %>% 
  cbind(as.matrix(bird_glm)[, "Stratum3"])
colnames(stratum3_estimates) <- c("bird_stglmm", "bird_glm")

(stratum3_plot <- stratum3_estimates %>% 
  mcmc_areas(pars = c("bird_stglmm", "bird_glm"), 
             prob = 0.95, 
             point_est = "median") +
  theme_minimal() +
  scale_y_discrete(labels = c("STGLMM", "GLM")))
ggsave("figures/3-20_bird-stratum3.png", stratum3_plot, width = 8, height = 6)
```

```{r}
cor_var_plot <- bird_stglmm$samples[5001:10000,,] %>% 
  mcmc_areas(pars = c("sigma2_sp", "sigma2_tm"),
             prob = 0.95) +
  scale_y_discrete(labels = unname(TeX(c("$\\sigma^2$", "$\\theta^2")))) +
  theme_minimal()

ggsave("figures/3-20_cor_var_plot.png", cor_var_plot, width = 8, height = 6)

```

